# Initialise Project
```{r load_project}
library(ProjectTemplate); load.project()
```             


# installation notes

Download and install R and Rstudio

* R: http://cran.rstudio.com/
* Rstudio: http://www.rstudio.com/products/RStudio/

Install lavaan and any other necessary packages

```{r}
# install.packages("lavaan")

# I also make us of 
# install.packages("psych")
# install.packages("ProjectTemplate")
```


# Help files

* http://lavaan.ugent.be/
The lavaan website is very user friendly.

* http://lavaan.ugent.be/tutorial/index.html
The site has many useful tutorials.

* R Help for lavaan is also very useful.
    * ?inspect : Help on the many different objects that can be extracted from a model fit
    * ?lavOptions : Help on the many additional arguments that can be used when model fitting with cfa/sem/growth/lavaan
    * ?"lavaan-class" : Shows the various methods that can be used with model fits
    
Original reference
Yves Rosseel (2012). lavaan: An R Package for Structural Equation Modeling. Journal of Statistical Software, 48(2), 1-36. URL http://www.jstatsoft.org/v48/i02/

* My lecture on CFA:
https://www.youtube.com/watch?v=BZWpAwd_xDs

* Video of CFA with lavaan:
https://www.youtube.com/watch?v=gcrXP2yMYY8



```{r}
# Load lavaan
library(lavaan)

# load data (e.g., from spss)
library(foreign)
ccases <- read.spss("data/ccases.sav", to.data.frame = TRUE, use.value.labels = FALSE)

# check data
head(ccases)
names(ccases) <- tolower(names(ccases))

# Model syntax http://lavaan.ugent.be/tutorial/syntax1.html
# ~ predict
# =~ indicator
# ~~ correlate
# ~1 intercept


# Step 1. Specify model
model <- "
agreeableness =~ a1 + a2 + a3
conscientiousness =~ c1 + c2 + c3
extraversion =~ e1 + e2 + e3
"

# View model
cat(model)

# fit model using CFA defaults
fit <- cfa(model = model, data = ccases)
# see also
# sem(model, ccases)

# viewing model
# overview of fit and parameters
summary(fit)

# Other arguments for summary method
# ?`lavaan-class`
# (object = "lavaan", header = TRUE, fit.measures=FALSE, estimates = TRUE, 
# ci = FALSE, fmi = FALSE, standardized = FALSE, 
# rsquare=FALSE, std.nox = FALSE, modindices=FALSE, ci=FALSE, nd = 3L)

summary(fit, standardized = TRUE)


# unstandardized parameters
parameterestimates(fit)

# standardized parameters
standardizedsolution(fit)

# Extract subset of parameters
# This is useful when you have many parameters
sfit <- standardizedsolution(fit)

# Just the loadings
sfit[ sfit$op == "=~", ] 

# Just correlations
sfit[ sfit$op == "~~" & sfit$lhs != sfit$rhs, ] 

# Just residual variances
sfit[ sfit$op == "~~" & sfit$lhs == sfit$rhs, ] 



########################
# fit measures
# all fit measures
fitmeasures(fit)

# Or select the subset of interest
fitmeasures(fit, c("npar", "chisq", "df", "cfi", "rmsea", "srmr"))



########################
# Other model features

# Look up the help for inspect to see many objects that can be returned
# ?inspect
# inspect(fit, what = "")

# variance explained
inspect(fit, 'r2')

# Sample means and covariance matrix
inspect(fit, "sampstat")

# Unstandardized model matrices
est_fit <- inspect(fit, "est")
est_fit$lambda # unstandardised loadings
est_fit$psi # latent variable covariance matrix

# Standardized model matrices
std_fit <- inspect(fit, "std")
std_fit$lambda # standardised loadings
std_fit$psi # latent variable correlation matrix


# compare models
fits <- list()
fits$fit1 <- fit

# orthogonal model
fits$fit0 <- cfa(model, ccases, orthogonal = TRUE)

# add residual 
# example of adding terms
# paste0 combines text
# "\n" represents new line
model2 <- paste0(model, "\n", "a1 ~~ a2")
# cat(model2)
fits$fit2 <- cfa(model2, ccases)

# combine into table
round(sapply(fits, fitmeasures), 3) # all output
# subset of fits
round(sapply(fits, function(X) 
    fitmeasures(X,  c("chisq", "df", "cfi", "rmsea", "srmr"))), 3)

# chisquare difference test
anova(fits$fit0, fits$fit1)


# Residuals
# unstandardized residuals 
# e.g., covariance matrix
resid(fit)

# standardized residuals
# 
resid(fit, type = "cor")

resid(fit, type = "cor")


# modification indices
mod_ind <- modificationindices(fit)

# you will get many modification indices
# and some will not make sense
# mod_ind

# top 10 softed by size
head(mod_ind[order(mod_ind$mi, decreasing=TRUE), ], 10)

# all those above 10
ordered_mod <- mod_ind[order(mod_ind$mi, decreasing=TRUE), ]
ordered_mod[ordered_mod$mi > 10, ]


#######################
# add predicted values to data file
pfit <- data.frame(predict(fit))
ccases2 <- cbind(ccases, pfit)
# head(ccases2)



# sem models
ccases$male <- as.numeric(ccases$gender == 1)

sem_model <- "
agreeableness =~ a1 + a2 + a3
conscientiousness =~ c1 + c2 + c3
extraversion =~ e1 + e2 + e3

conscientiousness ~ age + male
agreeableness ~ age + male
extraversion  ~ age + male
"

semfit <- sem(sem_model, ccases)

pfit <- parameterestimates(semfit)
pfit[pfit$op == "~", ]

sfit <- standardizedsolution(semfit)
sfit[sfit$op == "~", ]
```


# Plotting
```{r}
# install.packages("semPlot")
library(semPlot)
semPaths(semfit, "std")
```


